Acceleration of plague outbreaks in the second pandemic\\\emph{Electronic \supp}

### 2020-06-12 supp.Rnw
amsync

In the idealized situation in which the generation interval is exponentially distributed (as in the standard SIR model) then \cite{WallLips07}, as mentioned in the Discussion in the main text, 
%
\begin{linenomath*}
\begin{equation}\label{E:rofR}
r = \frac{\R_0-1}\gentime \,,
\end{equation}
\end{linenomath*}
%
where $\gentime$ is the mean generation interval.  Consequently, if this simple relationship holds and the mean generation interval \emph{does not change}, then a change in growth rate $r_1\to r_2$ implies a change in basic reproduction number $\R_{0,1}\to\R_{0,2}$, where
%
\begin{linenomath*}
\begin{equation}\label{E:1to2}
\frac{\R_{0,2}}{\R_{0,1}} = \frac{r_2\gentime+1}{r_1\gentime+1} \,.
\end{equation}
\end{linenomath*}
%
If we assume the estimate of $\gentime=18$ days obtained in
\cref{sec:gentime} above, and the MLE growth rates from early and late wills
listed in \cref{tab:params}, the relative change in reproduction
number that needs explanation is
<<rat_R0_calc,echo=FALSE>>=
ratR0 <- function(r){ (r / 365 * 18) + 1 }
load("analysis/fits/.epochsum.RData")
r1 <- rvals["fit","early"]
r2 <- rvals["fit","late"]
R01 <- ratR0(r1)
R02 <- ratR0(r2)
R0ratio <- R02/R01
@ 
%
\begin{linenomath*}
\begin{equation}\label{E:Rratio}
\frac{\R_{0,2}}{\R_{0,1}} 
\approx
\frac
{\Sexpr{round(r2,1)} \cdot \frac{18}{365} + 1}
{ \Sexpr{round(r1,2)} \cdot \frac{18}{365} + 1}
\approx
\frac{\Sexpr{round(R02,2)}}{\Sexpr{round(R01,2)}} 
\approx \Sexpr{round(R0ratio,2)}
\,.
\end{equation}
\end{linenomath*}
%
where the second subscript (1 or 2) denotes the early or late epoch, respectively.

>>>>

In the idealized situation in which the generation interval is exponentially distributed (as in the standard SIR model) then \cite{WallLips07}, as mentioned in the Discussion in the main text, 
%
\begin{linenomath*}
\begin{equation}\label{E:rofR}
r = \frac{\R_0-1}\gentime \,,
\end{equation}
\end{linenomath*}
%
where $\gentime$ is the mean generation interval.  Consequently, if this simple relationship holds and the mean generation interval \emph{does not change}, then a change in growth rate $r_1\to r_2$ implies a change in basic reproduction number $\R_{0,1}\to\R_{0,2}$, where
%
\begin{linenomath*}
\begin{equation}\label{E:1to2}
\frac{\R_{0,2}}{\R_{0,1}} = \frac{r_2\gentime+1}{r_1\gentime+1} \,.
\end{equation}
\end{linenomath*}
%
If we assume the estimate of $\gentime=18$ days obtained in
\cref{sec:gentime} above, and the MLE growth rates from early and late wills
listed in \cref{tab:params}, the relative change in reproduction
number that needs explanation is
<<rat_R0_calc,echo=FALSE>>=
ratR0 <- function(r){ (r / 365 * 18) + 1 }
load("analysis/fits/epochsum.RData")
r1 <- rvals["fit","early"]
r2 <- rvals["fit","late"]
R01 <- ratR0(r1)
R02 <- ratR0(r2)
R0ratio <- R02/R01
@ 
%
\begin{linenomath*}
\begin{equation}\label{E:Rratio}
\frac{\R_{0,2}}{\R_{0,1}} 
\approx
\frac
{\Sexpr{round(r2,1)} \cdot \frac{18}{365} + 1}
{ \Sexpr{round(r1,2)} \cdot \frac{18}{365} + 1}
\approx
\frac{\Sexpr{round(R02,2)}}{\Sexpr{round(R01,2)}} 
\approx \Sexpr{round(R0ratio,2)}
\,.
\end{equation}
\end{linenomath*}
%
where the second subscript (1 or 2) denotes the early or late epoch, respectively.
